\documentclass[12 pt]{article}
\renewcommand{\baselinestretch}{1}
\usepackage{geometry}
\usepackage{graphics}
\geometry{verbose,letterpaper,tmargin=0.5 cm,bmargin=0.8 cm,lmargin=2.5cm,rmargin=2.5cm,headsep=1cm}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{10 pt}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{setspace}
\usepackage{indentfirst}
\usepackage{hyperref}
\title{Predictive analytics with online change detection in data streams}
\date{}
\begin{document}
	\maketitle
  \section{SINE1 example}
  But concept drift (CD) can happen and the model performance can drop. In
  order to detect CD, change detectors are used to monitor changes in models
  performance metrics on-line. Once change is alarmed the model can be
  re-trained using the most relevant data between last change point and current
  time moment.

  Streaming settings is when the model is re-trained using new data not present
  in the initial training data set. Model is re-trained using all available
  data after each arrival of the new portion of data.  

  Figure XXX illustrates
  CD phenomena when SVM model is applied in streaming settings to the SINE1
  dataset. Figure XXX shows decision boundary before and after CD in SINE1
  dataset. Performance drops after moment $t$. And detectors alarms the change
  at the moment $t_1$.
  
  Before concept drift error rate decreases according to the statistical decision theory cite{XXX} as model is adaptively retrained.

  Figure XXX illustrates an example of adaptive learning with detector and
  without. It can be seen that with detector performance is better and with
  recurrent detector it is even more better because we can skip possible FA and
  we can possibly reduce the detection delay.

\end{document}
